---
title: "phase-chromosomes-snakeflow"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is just a little workflow I put together for phasing genotypes in a vcf.gz file
using `eagle` (and eventually I will add BEAGLE 4 in here for low coverage data, using 
genotype likelihoods).

The basic idea is that you give it:

1. The path to a big vcf.gz of bcf file (it must have an index, .tbi or .csi)
with all the genotypes,
2. a white-space delimited file where the first column holds the
chromosome names (corresponding to what you have in the VCF file)
and the second holds the integer equivalent, for the chromosomes
that you want to phase.  For example if we wanted to do the first
7 chromosomes of _O. mykiss_ that file would look like:
```
omy01 1
omy02 2
omy03 3
omy04 4
omy05 5
omy06 6
omy07 7
```

The steps here are to:

1. Break the VCF file up into a bunch of smaller BCF files, one per chromosome that you
want to phase, and the chromosomes in each to use simple integers. Then index each of
those.  These go into the `resources/rcBCF` directory.
2. Launch a separate job to phase each chromosome, by default using 20 threads.
3. In the end, the phased BCF file for each chromosome gets indexed by bcftools.


## Configuration

Currently, you can set what you need in the `config/config.yaml` file.  That looks
like this:
```yaml
`r paste(readLines("config/config.yaml"), collapse = "\n")`
```

## Output files

The phased BCF files and their indexes are produced with names/paths like:
```
results/phased_bcf/chr-1.bcf
results/phased_bcf/chr-1.bcf.csi
results/phased_bcf/chr-2.bcf
results/phased_bcf/chr-2.bcf.csi
results/phased_bcf/chr-3.bcf
results/phased_bcf/chr-3.bcf.csi
...
```

Like a standard snakemake workflow, the log files appear in 
`results/log/rule-name` where `rule-name` is the name of the
rule that created the log files.

## Command line invocation

**For a dry-run**
```sh
snakemake -np
```

**For running it on a node with 20 cores**
```sh
snakemake --cores 20 --use-conda -p
```
Note that the phasing is done, by default with 20 threads per job/chromosome.
If you need to change that, edit the Snakefile.

**For running it across multiple SLURM nodes each with 20 cores and 94Gb of RAM**
```
snakemake --jobs 100  -p --use-conda --profile sedna_slurm_profile/
```

## Rule Graph

Here is the rulegraph for this simple workflow, made with the command:
```sh
snakemake --rulegraph | dot -Tsvg > figs/rulegraph.svg
```
```{r, echo = FALSE, fig.align='center', out.width="25%"}
knitr::include_graphics("figs/rulegraph.svg")
```
